---
title: "explore_STRAPP_test_types"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{explore_STRAPP_test_types}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r set_options, include = FALSE}
knitr::opts_chunk$set(
  eval = FALSE, # Chunks of codes will not be evaluated by default
  collapse = TRUE,
  comment = "#>",
  fig.width = 7, fig.height = 5   # Set device size at rendering time (when plots are generated)
)
```

```{r setup, eval = TRUE, include = FALSE}
library(deepSTRAPP)
```

<br>
This vignette presents the different __types of STRAPP tests__ that can be carried on with deepSTRAPP across the different types of data (i.e., continuous, binary, and multinominal).

<br>
Two-tailed tests make no assumption on the direction of the effect tested. 

One-tailed tests make an hypothesis on the direction of the effect tested.

* For continuous data, the direction of the correlation is either __"positive"__ or __"negative"__.
* For binary data, the rate difference between state A and state B can be either __"A > B"__ or __"B > A"__.
* For multinominal data, overall Kruskal-Wallis tests are one-tailed by definition. They test for a rate difference across all states/ranges at once.
* Post hoc pairwise Dunn's tests carried out between pairs of states/ranges can be one-tailed or two-tailed, with all configuration tested by default.

<br>
Please note that the trait data and phylogeny calibration used in these examples are __NOT valid biological data__. They were modified in order to provide results illustrating the usefulness of deepSTRAPP.

<br>

```{r STRAPP_tests_cont_two_tailed}
# ------ Example 1: Continuous trait data ------ #

#### Step 1: Prepare data ####

## Load phylogeny with old time-calibration
data("Ponerinae_tree_old_calib", package = "deepSTRAPP")
## Load trait df
data("Ponerinae_trait_tip_data", package = "deepSTRAPP")

# Extract continuous trait data as a named vector
Ponerinae_cont_tip_data <- setNames(object = Ponerinae_trait_tip_data$fake_cont_tip_data,
                                    nm = Ponerinae_trait_tip_data$Taxa)
# Reorder tip data as in phylogeny
Ponerinae_cont_tip_data <- Ponerinae_cont_tip_data[Ponerinae_tree_old_calib$tip.label]

# Run prepare_trait_data with default options
# For continuous trait, a BM model is assumed by default.
Ponerinae_trait_object <- prepare_trait_data(tip_data = Ponerinae_cont_tip_data,
                                             trait_data_type = "continuous",
                                             phylo = Ponerinae_tree_old_calib,
                                             evolutionary_model = "BM",
                                             seed = 1234) # Set seed for reproducibility
                                              

#### Step 2: Run a two-tailed test ####

# No assumption is made on the direction of the effect tested.
# The null hypothesis is that no correlation can be found between rates and trait values.
# The alternative hypothesis is that there is a correlation between rates and trait values,
# in any direction (i.e., "positive" or "negative").

# Here, we run a two-tailed STRAPP test for t = 10 Mya.
focal_time <- 10

deepSTRAPP_output_two_tailed <- run_deepSTRAPP_for_focal_time(
   contMap = Ponerinae_trait_object$contMap,
   ace = Ponerinae_trait_object$ace,
   tip_data = Ponerinae_cont_tip_data,
   trait_data_type = "continuous",
   BAMM_object = Ponerinae_BAMM_object_old_calib,
   focal_time = focal_time,
   two_tailed = TRUE, # Select a two-tailed test (Default)
   rate_type = "net_diversification",
   seed = 1234, # Set seed for reproducibility
   return_perm_data = TRUE, # Keep permutation data to be able to plot histograms of tests
   # Needed to access traits and rates data to evaluate the direction of the correlation
   extract_diversification_data_melted_df = TRUE, 
   return_updated_trait_data_with_Map = TRUE) 
 
## Explore output
str(deepSTRAPP_output_two_tailed, max.level = 1)

# Access deepSTRAPP results
deepSTRAPP_output_two_tailed$STRAPP_results[1:5]

# We obtain a p-value = 0.016 leading us to discard the null hypothesis (no correlation).
# The estimate is the quantile 5% of the distribution of differences in 
# absolute Spearman rho-stats between observed and permuted data.
# The test is significant as the estimate is higher than zero.
# However, without plotting the rates vs. trait data, we do not know the direction of this correlation.

# Plot histogram of Spearman sum-rank test stats
plot_histogram_STRAPP_test_for_focal_time(
  deepSTRAPP_outputs = deepSTRAPP_output_two_tailed)

# The test is significant as the red line (quantile 5% = 0.063) is above the null expectation 
# that Δ abs(Spearman rho-stats) = 0.
```
```{r STRAPP_tests_cont_two_tailed_eval, fig.width = 8.5, fig.height = 6, out.width = "100%", eval = TRUE, echo = FALSE}
## Load phylogeny with old time-calibration
data("Ponerinae_tree_old_calib", package = "deepSTRAPP")
## Load trait df
data("Ponerinae_trait_tip_data", package = "deepSTRAPP")

# Extract continuous trait data as a named vector
Ponerinae_cont_tip_data <- setNames(object = Ponerinae_trait_tip_data$fake_cont_tip_data,
                                    nm = Ponerinae_trait_tip_data$Taxa)
# Reorder tip data as in phylogeny
Ponerinae_cont_tip_data <- Ponerinae_cont_tip_data[Ponerinae_tree_old_calib$tip.label]

# Run prepare_trait_data with default options
Ponerinae_trait_object <- suppressWarnings(prepare_trait_data(
  tip_data = Ponerinae_cont_tip_data,
  trait_data_type = "continuous",
  phylo = Ponerinae_tree_old_calib,
  evolutionary_model = "BM",
  seed = 1234,# Set seed for reproducibility
  plot_map = FALSE,
  verbose = FALSE)) 

# Here, we run a two-tailed STRAPP test for t = 10 Mya.
deepSTRAPP_output_two_tailed <- suppressWarnings(run_deepSTRAPP_for_focal_time(
   contMap = Ponerinae_trait_object$contMap,
   ace = Ponerinae_trait_object$ace,
   tip_data = Ponerinae_cont_tip_data,
   trait_data_type = "continuous",
   BAMM_object = Ponerinae_BAMM_object_old_calib,
   focal_time = 10,
   two_tailed = TRUE, # Select a two-tailed test (Default)
   rate_type = "net_diversification",
   seed = 1234, # Set seed for reproducibility
   return_perm_data = TRUE, # Keep permutation data to be able to plot histograms of tests
   # Needed to access traits and rates data to evaluate the direction of the correlation
   extract_diversification_data_melted_df = TRUE, 
   return_updated_trait_data_with_Map = TRUE,
   verbose = FALSE))

# Plot histogram of Spearman sum-rank test stats
plot_histogram_STRAPP_test_for_focal_time(
   deepSTRAPP_outputs = deepSTRAPP_output_two_tailed)
```

```{r STRAPP_tests_cont_two_tailed_2}
## Plot rates vs. trait values

# Extract mean diversification data, only for net diversification rates
diversification_data_df <- deepSTRAPP_output_two_tailed$diversification_data_df |>
  dplyr::filter(rate_type == "net_diversification") |>
  dplyr::group_by(tip_ID) |>
  dplyr::summarize(mean_rates = mean(rates))

# Extract trait data
trait_data_df <- as.data.frame(deepSTRAPP_output_two_tailed$updated_trait_data_with_Map$trait_data)
trait_data_df$tip_ID <- names(deepSTRAPP_output_two_tailed$updated_trait_data_with_Map$trait_data)
names(trait_data_df) <- c("trait_values", "tip_ID")

# Merge rates and traits in a unique data.frame
data_melted_df <- dplyr::left_join(x = diversification_data_df, y = trait_data_df,
                                   by = dplyr::join_by(tip_ID))

# Plot rates vs. trait values across all BAMM samples
ggplot_rates_vs_traits <- ggplot2::ggplot(data = data_melted_df) +
  # Plot points for all samples
  ggplot2::geom_point(mapping = ggplot2::aes(y = mean_rates, x = trait_values),
                      alpha = 0.80, size = 3) +
  # Adjust axis titles
  ggplot2::ylab("Net diversification rates") +
  ggplot2::xlab("Trait values") +
  
  # Adjust aesthetics
  ggplot2::theme(
     plot.margin = ggplot2::margin(0.3, 0.5, 0.5, 0.5, "inches"), # trbl
     panel.grid.major = ggplot2::element_line(color = "grey70", linetype = "dashed", linewidth = 0.3),
     panel.background = ggplot2::element_rect(fill = NA, color = NA),
     axis.title = ggplot2::element_text(size = 20, color = "black"),
     axis.title.x = ggplot2::element_text(margin = ggplot2::margin(t = 10)),
     axis.title.y = ggplot2::element_text(margin = ggplot2::margin(r = 12)),
     axis.line = ggplot2::element_line(linewidth = 1.0),
     axis.text = ggplot2::element_text(size = 18, color = "black"))
  
print(ggplot_rates_vs_traits)

# The plot indicates that rates are globally negatively correlated with trait values. 
# We could also have directly tested for hypothesis with a one-tailed test (See Step 3 below).

```
```{r STRAPP_tests_cont_two_tailed_eval_2, fig.width = 8.5, fig.height = 6, out.width = "100%", eval = TRUE, echo = FALSE}

## Plot rates vs. trait values

# Extract mean diversification data, only for net diversification rates
diversification_data_df <- deepSTRAPP_output_two_tailed$diversification_data_df |>
  dplyr::filter(rate_type == "net_diversification") |>
  dplyr::group_by(tip_ID) |>
  dplyr::summarize(mean_rates = mean(rates))

# Extract trait data
trait_data_df <- as.data.frame(deepSTRAPP_output_two_tailed$updated_trait_data_with_Map$trait_data)
trait_data_df$tip_ID <- names(deepSTRAPP_output_two_tailed$updated_trait_data_with_Map$trait_data)
names(trait_data_df) <- c("trait_values", "tip_ID")

# Merge rates and traits in a unique data.frame
data_melted_df <- dplyr::left_join(x = diversification_data_df, y = trait_data_df, by = dplyr::join_by(tip_ID))

# Plot rates vs. trait values across all BAMM samples
ggplot_rates_vs_traits <- ggplot2::ggplot(data = data_melted_df) +
  # Plot points for all samples
  ggplot2::geom_point(mapping = ggplot2::aes(y = mean_rates, x = trait_values),
                      alpha = 0.80, size = 3) +
  # Adjust axis titles
  ggplot2::ylab("Net diversification rates") +
  ggplot2::xlab("Trait values") +
  
  # Adjust aesthetics
  ggplot2::theme(
     plot.margin = ggplot2::margin(0.3, 0.5, 0.5, 0.5, "inches"), # trbl
     panel.grid.major = ggplot2::element_line(color = "grey70", linetype = "dashed", linewidth = 0.3),
     panel.background = ggplot2::element_rect(fill = NA, color = NA),
     axis.title = ggplot2::element_text(size = 20, color = "black"),
     axis.title.x = ggplot2::element_text(margin = ggplot2::margin(t = 10)),
     axis.title.y = ggplot2::element_text(margin = ggplot2::margin(r = 12)),
     axis.line = ggplot2::element_line(linewidth = 1.0),
     axis.text = ggplot2::element_text(size = 18, color = "black"))
  
print(ggplot_rates_vs_traits)

```

```{r STRAPP_tests_cont_one_tailed}
#### Step 3: Run a one-tailed test ####

# Here, we make an assumption on the direction of the effect tested.
# The null hypothesis is that there is a positive or no correlation between rates and trait values.
# The alternative hypothesis is that there is a negative correlation between rates and trait values.

# We run a one-tailed STRAPP test for t = 10 Mya.
focal_time <- 10

deepSTRAPP_output_one_tailed <- run_deepSTRAPP_for_focal_time(
   contMap = Ponerinae_trait_object$contMap,
   ace = Ponerinae_trait_object$ace,
   tip_data = Ponerinae_cont_tip_data,
   trait_data_type = "continuous",
   BAMM_object = Ponerinae_BAMM_object_old_calib,
   focal_time = focal_time,
   two_tailed = FALSE, # Select a one-tailed test
   one_tailed_hypothesis = "negative", # We select a "negative" correlation as the alternative hypothesis
   rate_type = "net_diversification",
   seed = 1234, # Set seed for reproducibility
   return_perm_data = TRUE) # Keep permutation data to be able to plot histograms of tests
 
## Explore output
str(deepSTRAPP_output_one_tailed, max.level = 1)

# Access deepSTRAPP results
deepSTRAPP_output_one_tailed$STRAPP_results[1:5]

# We obtain a p-value = 0.006 leading us to discard the null hypothesis
# (no correlation or positive correlation).
# Thus, the alternative hypothesis (negative correlation) is supported by the test.
# The estimate is the quantile 95% of the distribution of differences in Spearman rho-stats
# between observed and permuted data.
# The test is significant as the estimate is lower than zero.

# Plot histogram of Spearman sum-rank test stats
plot_histogram_STRAPP_test_for_focal_time(
  deepSTRAPP_outputs = deepSTRAPP_output_one_tailed)

# The test is significant as the red line (quantile 95% = -0.147) is below the null expectation 
# that Δ Spearman rho stat = 0.

```
```{r STRAPP_tests_cont_one_tailed_eval, fig.width = 8.5, fig.height = 6, out.width = "100%", eval = TRUE, echo = FALSE}
# We run a one-tailed STRAPP test for t = 10 Mya.
deepSTRAPP_output_one_tailed <- suppressWarnings(run_deepSTRAPP_for_focal_time(
   contMap = Ponerinae_trait_object$contMap,
   ace = Ponerinae_trait_object$ace,
   tip_data = Ponerinae_cont_tip_data,
   trait_data_type = "continuous",
   BAMM_object = Ponerinae_BAMM_object_old_calib,
   focal_time = 10,
   two_tailed = FALSE, # Select a one-tailed test
   one_tailed_hypothesis = "negative", # We select a "negative" correlation as the alternative hypothesis
   rate_type = "net_diversification",
   seed = 1234, # Set seed for reproducibility
   return_perm_data = TRUE, # Keep permutation data to be able to plot histograms of tests
   verbose = FALSE))

# Plot histogram of Spearman sum-rank test stats
plot_histogram_STRAPP_test_for_focal_time(
  deepSTRAPP_outputs = deepSTRAPP_output_one_tailed)
```


```{r STRAPP_tests_cat_2lvl_two_tailed}
# ------ Example 2: Categorical binary trait data ------ #

#### Step 1: Prepare data ####

## Load phylogeny with old time-calibration
data("Ponerinae_tree_old_calib", package = "deepSTRAPP")
## Load trait df
data("Ponerinae_trait_tip_data", package = "deepSTRAPP")

# Extract continuous trait data as a named vector
Ponerinae_cat_2lvl_tip_data <- setNames(object = Ponerinae_trait_tip_data$fake_cat_2lvl_tip_data,
                                    nm = Ponerinae_trait_tip_data$Taxa)
# Reorder tip data as in phylogeny
Ponerinae_cat_2lvl_tip_data <- Ponerinae_cat_2lvl_tip_data[Ponerinae_tree_old_calib$tip.label]

# Select color scheme for states
colors_per_states <- c("darkblue", "lightblue")
names(colors_per_states) <- c("large", "small")

# Run prepare_trait_data with default options
Ponerinae_trait_object <- prepare_trait_data(tip_data = Ponerinae_cat_2lvl_tip_data,
                                             trait_data_type = "categorical",
                                             phylo = Ponerinae_tree_old_calib,
                                             evolutionary_model = "ARD", # Select an Equal-Rates model
                                             colors_per_states = colors_per_states,
                                             nb_simulations = 100,
                                             seed = 1234) # Set seed for reproducibility
# Load directly output to save time
data(Ponerinae_cat_2lvl_data_old_calib, package = "deepSTRAPP")
Ponerinae_trait_object <- Ponerinae_cat_2lvl_data_old_calib


#### Step 2: Run a two-tailed test ####

# No assumption is made ab out which states exhibit the highest rates.
# The null hypothesis is that there is no difference in rates between the two states.
# The alternative hypothesis is that there is a difference in rates between the two states,
# in any direction (i.e., "small > large" or "large > small").

# Here, we run a two-tailed STRAPP test for t = 10 Mya.
focal_time <- 10

deepSTRAPP_output_two_tailed <- run_deepSTRAPP_for_focal_time(
   densityMaps = Ponerinae_trait_object$densityMaps,
   ace = Ponerinae_trait_object$ace,
   tip_data = Ponerinae_cat_2lvl_tip_data,
   trait_data_type = "categorical",
   BAMM_object = Ponerinae_BAMM_object_old_calib,
   focal_time = focal_time,
   two_tailed = TRUE, # Select a two-tailed test (Default)
   rate_type = "net_diversification",
   seed = 1234, # Set seed for reproducibility
   return_perm_data = TRUE, # Keep permutation data to be able to plot histograms of tests
   # Needed to access traits and rates data to evaluate the direction of the correlation
   extract_diversification_data_melted_df = TRUE, 
   return_updated_trait_data_with_Map = TRUE) 

## Explore output
str(deepSTRAPP_output_two_tailed, max.level = 1)

# Access deepSTRAPP results
deepSTRAPP_output_two_tailed$STRAPP_results[1:5]

# We obtain a p-value = 0.037 leading us to discard the null hypothesis (no differences).
# The estimate is the quantile 5% of the distribution of differences 
# in absolute Mann-Whitney-Wilcoxon U-stats between observed and permuted data.
# The test is significant as the estimate is higher than zero.
# However, without plotting the rates vs. states, we do not know the direction of this difference.

# Plot histogram of Mann-Whitney-Wilcoxon test stats
plot_histogram_STRAPP_test_for_focal_time(
  deepSTRAPP_outputs = deepSTRAPP_output_two_tailed)

# The test is significant as the red line (quantile 5% = 463.6) is above the null expectation 
# that Δ abs(Mann-Whitney-Wilcoxon U-stats) = 0.
```
```{r STRAPP_tests_cat_2lvl_two_tailed_eval, fig.width = 8.5, fig.height = 6, out.width = "100%", eval = TRUE, echo = FALSE}
# Load phylogeny with old time-calibration
data("Ponerinae_tree_old_calib", package = "deepSTRAPP")
# Load trait df
data("Ponerinae_trait_tip_data", package = "deepSTRAPP")

# Extract continuous trait data as a named vector
Ponerinae_cat_2lvl_tip_data <- setNames(object = Ponerinae_trait_tip_data$fake_cat_2lvl_tip_data,
                                    nm = Ponerinae_trait_tip_data$Taxa)
# Reorder tip data as in phylogeny
Ponerinae_cat_2lvl_tip_data <- Ponerinae_cat_2lvl_tip_data[Ponerinae_tree_old_calib$tip.label]

# Load directly deepSTRAPP output to save time
data(Ponerinae_cat_2lvl_data_old_calib, package = "deepSTRAPP")
Ponerinae_trait_object <- Ponerinae_cat_2lvl_data_old_calib

# Select color scheme for states
colors_per_states <- c("darkblue", "lightblue")
names(colors_per_states) <- c("large", "small")

# Here, we run a two-tailed STRAPP test for t = 10 Mya.
deepSTRAPP_output_two_tailed <- suppressWarnings(run_deepSTRAPP_for_focal_time(
   densityMaps = Ponerinae_trait_object$densityMaps,
   ace = Ponerinae_trait_object$ace,
   tip_data = Ponerinae_cat_2lvl_tip_data,
   trait_data_type = "categorical",
   BAMM_object = Ponerinae_BAMM_object_old_calib,
   focal_time = 10,
   two_tailed = TRUE, # Select a two-tailed test (Default)
   rate_type = "net_diversification",
   seed = 1234, # Set seed for reproducibility
   return_perm_data = TRUE, # Keep permutation data to be able to plot histograms of tests
   # Needed to access traits and rates data to evaluate the direction of the correlation
   extract_diversification_data_melted_df = TRUE, 
   return_updated_trait_data_with_Map = TRUE,
   verbose = FALSE)) 

# Plot histogram of Mann-Whitney-Wilcoxon test stats
plot_histogram_STRAPP_test_for_focal_time(
  deepSTRAPP_outputs = deepSTRAPP_output_two_tailed)
```

```{r STRAPP_tests_cat_2lvl_two_tailed_2}
## Plot rates vs. trait values

# Extract mean diversification data, only for net diversification rates
diversification_data_df <- deepSTRAPP_output_two_tailed$diversification_data_df |>
  dplyr::filter(rate_type == "net_diversification") |>
  dplyr::group_by(tip_ID) |>
  dplyr::summarize(mean_rates = mean(rates))

# Extract trait data
trait_data_df <- as.data.frame(deepSTRAPP_output_two_tailed$updated_trait_data_with_Map$trait_data)
trait_data_df$tip_ID <- names(deepSTRAPP_output_two_tailed$updated_trait_data_with_Map$trait_data)
names(trait_data_df) <- c("trait_values", "tip_ID")

# Merge rates and traits in a unique data.frame
data_melted_df <- dplyr::left_join(x = diversification_data_df, y = trait_data_df,
                                   by = dplyr::join_by(tip_ID))

# Plot rates vs. trait values across all BAMM samples
ggplot_rates_vs_traits <- ggplot2::ggplot(data = data_melted_df) +
  
  # Plot boxplot per states
  ggplot2::geom_boxplot(mapping = ggplot2::aes(y = mean_rates, x = trait_values,
                                               fill = trait_values)) +
  # Plot points for all samples
  ggplot2::geom_jitter(mapping = ggplot2::aes(y = mean_rates, x = trait_values,
                                              fill = trait_values),
                      alpha = 0.50, size = 3, width = 0.25, shape = 21, color = "black") +
  # Adjust axis titles
  ggplot2::ylab("Net diversification rates") +
  ggplot2::xlab("States") +
  
  # Adjust legend
  ggplot2::scale_fill_manual(name = "States",
                             breaks = names(colors_per_states),
                             values = colors_per_states) +
  
  # Adjust aesthetics
  ggplot2::theme(
     plot.margin = ggplot2::margin(0.3, 0.5, 0.5, 0.5, "inches"), # trbl
     panel.grid.major = ggplot2::element_line(color = "grey70", linetype = "dashed", linewidth = 0.3),
     panel.background = ggplot2::element_rect(fill = NA, color = NA),
     axis.title = ggplot2::element_text(size = 20, color = "black"),
     axis.title.x = ggplot2::element_text(margin = ggplot2::margin(t = 10)),
     axis.title.y = ggplot2::element_text(margin = ggplot2::margin(r = 12)),
     axis.line = ggplot2::element_line(linewidth = 1.0),
     axis.text = ggplot2::element_text(size = 18, color = "black"))
  
print(ggplot_rates_vs_traits)

# The plot indicates that net diversification rates are higher in "small" ants than in "large" ants. 
# We could also have directly tested for hypothesis with a one-tailed test (See Step 3 below).
 
```
```{r STRAPP_tests_cat_2lvl_two_tailed_eval2, fig.width = 8.5, fig.height = 6, out.width = "100%", eval = TRUE, echo = FALSE}
## Plot rates vs. trait values

# Extract mean diversification data, only for net diversification rates
diversification_data_df <- deepSTRAPP_output_two_tailed$diversification_data_df |>
  dplyr::filter(rate_type == "net_diversification") |>
  dplyr::group_by(tip_ID) |>
  dplyr::summarize(mean_rates = mean(rates))

# Extract trait data
trait_data_df <- as.data.frame(deepSTRAPP_output_two_tailed$updated_trait_data_with_Map$trait_data)
trait_data_df$tip_ID <- names(deepSTRAPP_output_two_tailed$updated_trait_data_with_Map$trait_data)
names(trait_data_df) <- c("trait_values", "tip_ID")

# Merge rates and traits in a unique data.frame
data_melted_df <- dplyr::left_join(x = diversification_data_df, y = trait_data_df, by = dplyr::join_by(tip_ID))

# Plot rates vs. trait values across all BAMM samples
ggplot_rates_vs_traits <- ggplot2::ggplot(data = data_melted_df) +
  
  # Plot boxplot per states
  ggplot2::geom_boxplot(mapping = ggplot2::aes(y = mean_rates, x = trait_values,
                                               fill = trait_values)) +
  # Plot points for all samples
  ggplot2::geom_jitter(mapping = ggplot2::aes(y = mean_rates, x = trait_values,
                                              fill = trait_values),
                      alpha = 0.50, size = 3, width = 0.25, shape = 21, color = "black") +
  # Adjust axis titles
  ggplot2::ylab("Net diversification rates") +
  ggplot2::xlab("States") +
  
  # Adjust legend
  ggplot2::scale_fill_manual(name = "States", breaks = names(colors_per_states), values = colors_per_states) +
  
  # Adjust aesthetics
  ggplot2::theme(
     plot.margin = ggplot2::margin(0.3, 0.5, 0.5, 0.5, "inches"), # trbl
     panel.grid.major = ggplot2::element_line(color = "grey70", linetype = "dashed", linewidth = 0.3),
     panel.background = ggplot2::element_rect(fill = NA, color = NA),
     axis.title = ggplot2::element_text(size = 20, color = "black"),
     axis.title.x = ggplot2::element_text(margin = ggplot2::margin(t = 10)),
     axis.title.y = ggplot2::element_text(margin = ggplot2::margin(r = 12)),
     axis.line = ggplot2::element_line(linewidth = 1.0),
     axis.text = ggplot2::element_text(size = 18, color = "black"))
  
print(ggplot_rates_vs_traits)

```

```{r STRAPP_tests_cat_2lvl_one_tailed}
#### Step 3: Run a one-tailed test ####

# Here, we make an assumption on the direction of the effect tested.
# The null hypothesis is that diversification rates are lower or equal
# for "small" ants compared to "large" ants.
# The alternative hypothesis is that diversification rates are higher
# for "small" ants than "large" ants.

# We run a one-tailed STRAPP test for t = 10 Mya.
focal_time <- 10

deepSTRAPP_output_one_tailed <- run_deepSTRAPP_for_focal_time(
   densityMaps = Ponerinae_trait_object$densityMaps,
   ace = Ponerinae_trait_object$ace,
   tip_data = Ponerinae_cat_2lvl_tip_data,
   trait_data_type = "categorical",
   BAMM_object = Ponerinae_BAMM_object_old_calib,
   focal_time = focal_time,
   two_tailed = FALSE, # Select a one-tailed test
   one_tailed_hypothesis = "small > large", # We select "small > large" as the alternative hypothesis
   rate_type = "net_diversification",
   seed = 1234, # Set seed for reproducibility
   return_perm_data = TRUE) # Keep permutation data to be able to plot histograms of tests
 
## Explore output
str(deepSTRAPP_output_one_tailed, max.level = 1)

# Access deepSTRAPP results
deepSTRAPP_output_one_tailed$STRAPP_results[1:5]

# We obtain a p-value = 0.019 leading us to discard the null hypothesis ("small <= large").
# Thus, the alternative hypothesis ("small > large") is supported by the test.
# The estimate is the quantile 5% of the distribution of differences in Mann-Whitney-Wilcoxon U-stats
# between observed and permuted data
# The test is significant as the estimate is higher than zero.

# Plot histogram of Mann-Whitney-Wilcoxon test stats
plot_histogram_STRAPP_test_for_focal_time(
  deepSTRAPP_outputs = deepSTRAPP_output_one_tailed)

# The test is significant as the red line (quantile 5% = 3899.8) is above the null expectation 
# that Δ Mann-Whitney-Wilcoxon U-stats = 0.

```
```{r STRAPP_tests_cat_2lvl_one_tailed_eval, fig.width = 8.5, fig.height = 6, out.width = "100%", eval = TRUE, echo = FALSE}
# We run a one-tailed STRAPP test for t = 10 Mya.
deepSTRAPP_output_one_tailed <- suppressWarnings(run_deepSTRAPP_for_focal_time(
   densityMaps = Ponerinae_trait_object$densityMaps,
   ace = Ponerinae_trait_object$ace,
   tip_data = Ponerinae_cat_2lvl_tip_data,
   trait_data_type = "categorical",
   BAMM_object = Ponerinae_BAMM_object_old_calib,
   focal_time = 10,
   two_tailed = FALSE, # Select a one-tailed test
   one_tailed_hypothesis = "small > large", # We select "small > large" as the alternative hypothesis
   rate_type = "net_diversification",
   seed = 1234, # Set seed for reproducibility
   return_perm_data = TRUE, # Keep permutation data to be able to plot histograms of tests
   verbose = FALSE))

# Plot histogram of Mann-Whitney-Wilcoxon test stats
plot_histogram_STRAPP_test_for_focal_time(
  deepSTRAPP_outputs = deepSTRAPP_output_one_tailed)
```



```{r STRAPP_tests_cat_3lvl_overall}
# ------ Example 3: Categorical multinominal (3-levels) trait data ------ #

#### Step 1: Prepare data ####
## Load phylogeny with old time-calibration
data("Ponerinae_tree_old_calib", package = "deepSTRAPP")
## Load trait df
data("Ponerinae_trait_tip_data", package = "deepSTRAPP")

# Extract continuous trait data as a named vector
Ponerinae_cat_3lvl_tip_data <- setNames(object = Ponerinae_trait_tip_data$fake_cat_3lvl_tip_data,
                                    nm = Ponerinae_trait_tip_data$Taxa)
# Reorder tip data as in phylogeny
Ponerinae_cat_3lvl_tip_data <- Ponerinae_cat_3lvl_tip_data[Ponerinae_tree_old_calib$tip.label]

## Select color scheme for states
colors_per_states <- c("forestgreen", "sienna", "goldenrod")
names(colors_per_states) <- c("arboreal", "subterranean", "terricolous")

# Run prepare_trait_data with default options
Ponerinae_trait_object <- prepare_trait_data(tip_data = Ponerinae_cat_3lvl_tip_data,
                                             trait_data_type = "categorical",
                                             phylo = Ponerinae_tree_old_calib,
                                             evolutionary_model = "ARD", # Select an Equal-Rates model
                                             colors_per_states = colors_per_states,
                                             nb_simulations = 100,
                                             seed = 1234) # Set seed for reproducibility
# Load directly output to save time
data(Ponerinae_cat_3lvl_data_old_calib, package = "deepSTRAPP")
Ponerinae_trait_object <- Ponerinae_cat_3lvl_data_old_calib

#### Step 2: Run the overall Kruskal-Wallis test ####

# For multinominal data (i.e., with more than two states/ranges), the overall test
# is by definition a one-tailed Kruskal-Wallis test for rate differences across all states/ranges.
# There is no two-tailed version of this test, thus deepSTRAPP assumes a one-tailed test by default.
# The null hypothesis is that there is no difference in rates across all states/ranges.
# The alternative hypothesis is that there is at least one pairs of states/ranges
# with differences in diversification rates. 

# To test for differences between pairs of states/ranges, you need to run post hoc pairwise tests.
# These Dunn's test can be one-tailed or two-tailed depending on the null hypotheses stated.
# See the next Steps 3 & 4 for post hoc pairwise tests.

# Here, we run an overall STRAPP test for t = 10 Mya.
focal_time <- 10

deepSTRAPP_output_overall <- run_deepSTRAPP_for_focal_time(
   densityMaps = Ponerinae_trait_object$densityMaps,
   ace = Ponerinae_trait_object$ace,
   tip_data = Ponerinae_cat_3lvl_tip_data,
   trait_data_type = "categorical",
   BAMM_object = Ponerinae_BAMM_object_old_calib,
   focal_time = focal_time,
   alpha = 0.10, # Adjust the significance threshold
   posthoc_pairwise_tests = FALSE, # To run the overall 
   rate_type = "net_diversification",
   seed = 1234, # Set seed for reproducibility
   return_perm_data = TRUE, # Keep permutation data to be able to plot histograms of tests
   # Needed to access traits and rates data to evaluate the direction of the correlation
   extract_diversification_data_melted_df = TRUE, 
   return_updated_trait_data_with_Map = TRUE) 

## Explore output
str(deepSTRAPP_output_overall, max.level = 1)

# Access deepSTRAPP results
deepSTRAPP_output_overall$STRAPP_results[1:5]

# We obtain a p-value = 0.071 leading us to discard the null hypothesis (no differences).
# The estimate is the quantile 10% of the distribution of differences 
# in Kruskal-Wallis H-stats between observed and permuted data.
# The test is significant as the estimate is higher than zero.
# However, without plotting the rates vs. states, or running post hoc pairwise tests,
# we do not know which pairs of states exhibit differences.

# Plot histogram of Kruskall-Wallis one-way ANOVA test stats
plot_histogram_STRAPP_test_for_focal_time(
  deepSTRAPP_outputs = deepSTRAPP_output_overall)

# The test is significant as the red line (quantile 10% = 6.942) is above the null expectation 
# that Δ Kruskal-Wallis H-stats = 0.
```
```{r STRAPP_tests_cat_3lvl_overall_eval, fig.width = 8.5, fig.height = 6, out.width = "100%", eval = TRUE, echo = FALSE}
# Load phylogeny with old time-calibration
data("Ponerinae_tree_old_calib", package = "deepSTRAPP")
# Load trait df
data("Ponerinae_trait_tip_data", package = "deepSTRAPP")

# Extract continuous trait data as a named vector
Ponerinae_cat_3lvl_tip_data <- setNames(object = Ponerinae_trait_tip_data$fake_cat_3lvl_tip_data,
                                    nm = Ponerinae_trait_tip_data$Taxa)
# Reorder tip data as in phylogeny
Ponerinae_cat_3lvl_tip_data <- Ponerinae_cat_3lvl_tip_data[Ponerinae_tree_old_calib$tip.label]

# Load directly deepSTRAPP output to save time
data(Ponerinae_cat_3lvl_data_old_calib, package = "deepSTRAPP")
Ponerinae_trait_object <- Ponerinae_cat_3lvl_data_old_calib

## Select color scheme for states
colors_per_states <- c("forestgreen", "sienna", "goldenrod")
names(colors_per_states) <- c("arboreal", "subterranean", "terricolous")

# Here, we run an overall STRAPP test for t = 10 Mya.
deepSTRAPP_output_overall <- suppressWarnings(run_deepSTRAPP_for_focal_time(
   densityMaps = Ponerinae_trait_object$densityMaps,
   ace = Ponerinae_trait_object$ace,
   tip_data = Ponerinae_cat_3lvl_tip_data,
   trait_data_type = "categorical",
   BAMM_object = Ponerinae_BAMM_object_old_calib,
   focal_time = 10,
   alpha = 0.10, # Adjust the significance threshold
   posthoc_pairwise_tests = FALSE, # To run the overall 
   rate_type = "net_diversification",
   seed = 1234, # Set seed for reproducibility
   return_perm_data = TRUE, # Keep permutation data to be able to plot histograms of tests
   # Needed to access traits and rates data to evaluate the direction of the correlation
   extract_diversification_data_melted_df = TRUE, 
   return_updated_trait_data_with_Map = TRUE,
   verbose = FALSE)) 

# Plot histogram of Kruskall-Wallis one-way ANOVA test stats
plot_histogram_STRAPP_test_for_focal_time(
  deepSTRAPP_outputs = deepSTRAPP_output_overall)
```

```{r STRAPP_tests_cat_3lvl_overall_2}
## Plot rates vs. trait values

# Extract mean diversification data, only for net diversification rates
diversification_data_df <- deepSTRAPP_output_overall$diversification_data_df |>
  dplyr::filter(rate_type == "net_diversification") |>
  dplyr::group_by(tip_ID) |>
  dplyr::summarize(mean_rates = mean(rates))

# Extract trait data
trait_data_df <- as.data.frame(deepSTRAPP_output_overall$updated_trait_data_with_Map$trait_data)
trait_data_df$tip_ID <- names(deepSTRAPP_output_overall$updated_trait_data_with_Map$trait_data)
names(trait_data_df) <- c("trait_values", "tip_ID")

# Merge rates and traits in a unique data.frame
data_melted_df <- dplyr::left_join(x = diversification_data_df, y = trait_data_df,
                                   by = dplyr::join_by(tip_ID))

# Plot rates vs. trait values across all BAMM samples
ggplot_rates_vs_traits <- ggplot2::ggplot(data = data_melted_df) +
  
  # Plot boxplot per states
  ggplot2::geom_boxplot(mapping = ggplot2::aes(y = mean_rates, x = trait_values,
                                               fill = trait_values)) +
  # Plot points for all samples
  ggplot2::geom_jitter(mapping = ggplot2::aes(y = mean_rates, x = trait_values,
                                              fill = trait_values),
                      alpha = 0.50, size = 3, width = 0.25, shape = 21, color = "black") +
  # Adjust axis titles
  ggplot2::ylab("Net diversification rates") +
  ggplot2::xlab("States") +
  
  # Adjust legend
  ggplot2::scale_fill_manual(name = "States", breaks = names(colors_per_states), values = colors_per_states) +
  
  # Adjust aesthetics
  ggplot2::theme(
     plot.margin = ggplot2::margin(0.3, 0.5, 0.5, 0.5, "inches"), # trbl
     panel.grid.major = ggplot2::element_line(color = "grey70", linetype = "dashed", linewidth = 0.3),
     panel.background = ggplot2::element_rect(fill = NA, color = NA),
     axis.title = ggplot2::element_text(size = 20, color = "black"),
     axis.title.x = ggplot2::element_text(margin = ggplot2::margin(t = 10)),
     axis.title.y = ggplot2::element_text(margin = ggplot2::margin(r = 12)),
     axis.line = ggplot2::element_line(linewidth = 1.0),
     axis.text = ggplot2::element_text(size = 18, color = "black"))
  
print(ggplot_rates_vs_traits)

# The plot indicates that net diversification rates are higher in "terricolous" ants, 
# then "subterranean" ants, and lower in "arboreal" ants. 
# We can explore those potential differences with post hoc pairwise tests,
# with both one-tailed hypothesis tests or two-tailed tests (See Steps 3 & 4 below).
 
```
```{r STRAPP_tests_cat_3lvl_overall_eval2, fig.width = 8.5, fig.height = 6, out.width = "100%", eval = TRUE, echo = FALSE}
## Plot rates vs. trait values

# Extract mean diversification data, only for net diversification rates
diversification_data_df <- deepSTRAPP_output_overall$diversification_data_df |>
  dplyr::filter(rate_type == "net_diversification") |>
  dplyr::group_by(tip_ID) |>
  dplyr::summarize(mean_rates = mean(rates))

# Extract trait data
trait_data_df <- as.data.frame(deepSTRAPP_output_overall$updated_trait_data_with_Map$trait_data)
trait_data_df$tip_ID <- names(deepSTRAPP_output_overall$updated_trait_data_with_Map$trait_data)
names(trait_data_df) <- c("trait_values", "tip_ID")

# Merge rates and traits in a unique data.frame
data_melted_df <- dplyr::left_join(x = diversification_data_df, y = trait_data_df, by = dplyr::join_by(tip_ID))

# Plot rates vs. trait values across all BAMM samples
ggplot_rates_vs_traits <- ggplot2::ggplot(data = data_melted_df) +
  
  # Plot boxplot per states
  ggplot2::geom_boxplot(mapping = ggplot2::aes(y = mean_rates, x = trait_values,
                                               fill = trait_values)) +
  # Plot points for all samples
  ggplot2::geom_jitter(mapping = ggplot2::aes(y = mean_rates, x = trait_values,
                                              fill = trait_values),
                      alpha = 0.50, size = 3, width = 0.25, shape = 21, color = "black") +
  # Adjust axis titles
  ggplot2::ylab("Net diversification rates") +
  ggplot2::xlab("States") +
  
  # Adjust legend
  ggplot2::scale_fill_manual(name = "States", breaks = names(colors_per_states), values = colors_per_states) +
  
  # Adjust aesthetics
  ggplot2::theme(
     plot.margin = ggplot2::margin(0.3, 0.5, 0.5, 0.5, "inches"), # trbl
     panel.grid.major = ggplot2::element_line(color = "grey70", linetype = "dashed", linewidth = 0.3),
     panel.background = ggplot2::element_rect(fill = NA, color = NA),
     axis.title = ggplot2::element_text(size = 20, color = "black"),
     axis.title.x = ggplot2::element_text(margin = ggplot2::margin(t = 10)),
     axis.title.y = ggplot2::element_text(margin = ggplot2::margin(r = 12)),
     axis.line = ggplot2::element_line(linewidth = 1.0),
     axis.text = ggplot2::element_text(size = 18, color = "black"))
  
print(ggplot_rates_vs_traits)

```

```{r STRAPP_tests_cat_3lvl_two_tailed_eval2}
#### Step 3: Run all post hoc Dunn's two-tailed tests ####

# Post hoc tests are pairwise tests for differences between pairs of states/ranges.
# They can be two-tailed, without assumption regarding which states exhibit higher rates than the others.
# or they can be one-tailed, with assumption on which states exhibit higher rates within each pair.

# deepSTRAPP runs post hoc pairwise tests across all pairs of states.
# You can choose which type of tests (i.e., "two-tailed" or "one-tailed") to run.
# Here, the two-tailed version is presented. See Step 4 for the one-tailed version.

# In the two-tailed tests, the null hypothesis for each pair is that there is 
# no difference in rates between the two states.
# The alternative hypothesis is that there is a difference in rates between the two states,
# in any direction (e.g., "arboreal > terricolous" or "terricolous > arboreal").

# You can also select a method to adjust p-values for multiple testing.
# See stats::p.adjust() for the available methods.
# Here, we select none.

# We run a two-tailed STRAPP test for t = 10 Mya.
focal_time <- 10

deepSTRAPP_output_two_tailed <- run_deepSTRAPP_for_focal_time(
   densityMaps = Ponerinae_trait_object$densityMaps,
   ace = Ponerinae_trait_object$ace,
   tip_data = Ponerinae_cat_3lvl_tip_data,
   trait_data_type = "categorical",
   BAMM_object = Ponerinae_BAMM_object_old_calib,
   focal_time = focal_time,
   two_tailed = TRUE, # Select two-tailed tests for the post hoc tests (Default)
   posthoc_pairwise_tests = TRUE, # Ask for post hoc tests to be carried out
   rate_type = "net_diversification",
   seed = 1234, # Set seed for reproducibility
   return_perm_data = TRUE) # Keep permutation data to be able to plot histograms of tests

## Explore output
str(deepSTRAPP_output_two_tailed, max.level = 2)

# Access deepSTRAPP results for post hoc tests
deepSTRAPP_output_two_tailed$STRAPP_results$posthoc_pairwise_tests$summary_df

# We obtain a p-value for each pair of states. Here there are 3 possible pairs.
# Only the pair "arboreal != terricolous" yields to a significant result.
# For this pair, we obtain a p-value = 0.025 leading us to discard the null hypothesis (no differences).
# The estimate is the quantile 5% of the distribution of differences 
# in absolute Dunn's Z-stats between observed and permuted data.
# The test is significant as the estimate is higher than zero (Q5% = 0.759).
# However, without plotting the rates vs. states, we would have not know the direction of this difference.

# Plot histogram of all post hoc two-tailed test stats
plot_histogram_STRAPP_test_for_focal_time(
  deepSTRAPP_outputs = deepSTRAPP_output_two_tailed,
  plot_posthoc_tests = TRUE)

# Only the test for the pair "arboreal =! terricolous" is significant as the red line
# (quantile 5% = 0.759) is above the null expectation that Δ Dunn's Z-stats = 0.
# The conclusion is that it is the difference in rates between "arboreal" ants
# and "terricolous" ants that drives the differences in rates recorded overall
# with the Kruskal-Wallis test.
# Yet, with this test only, we do not know which states has the higher rates
# We can look at the plot of rates vs. states above (See Step 2),
# but we can also test directly for the two hypotheses (i.e., ("arboreal" > "terricolous")
# or ("terricolous" > "arboreal") using one-tailed tests.
# See Step 4 below for the one-tailed tests.
```
```{r STRAPP_tests_cat_3lvl_two_tailed_eval, fig.width = 8.5, fig.height = 7, out.width = "100%", eval = TRUE, echo = FALSE}
# Load phylogeny with old time-calibration
data("Ponerinae_tree_old_calib", package = "deepSTRAPP")
# Load trait df
data("Ponerinae_trait_tip_data", package = "deepSTRAPP")

# Extract continuous trait data as a named vector
Ponerinae_cat_3lvl_tip_data <- setNames(object = Ponerinae_trait_tip_data$fake_cat_3lvl_tip_data,
                                    nm = Ponerinae_trait_tip_data$Taxa)
# Reorder tip data as in phylogeny
Ponerinae_cat_3lvl_tip_data <- Ponerinae_cat_3lvl_tip_data[Ponerinae_tree_old_calib$tip.label]

# Load directly deepSTRAPP output to save time
data(Ponerinae_cat_3lvl_data_old_calib, package = "deepSTRAPP")
Ponerinae_trait_object <- Ponerinae_cat_3lvl_data_old_calib

## Select color scheme for states
colors_per_states <- c("forestgreen", "sienna", "goldenrod")
names(colors_per_states) <- c("arboreal", "subterranean", "terricolous")

# Here, we run a two-tailed STRAPP test for t = 10 Mya.
deepSTRAPP_output_two_tailed <- suppressWarnings(run_deepSTRAPP_for_focal_time(
   densityMaps = Ponerinae_trait_object$densityMaps,
   ace = Ponerinae_trait_object$ace,
   tip_data = Ponerinae_cat_3lvl_tip_data,
   trait_data_type = "categorical",
   BAMM_object = Ponerinae_BAMM_object_old_calib,
   focal_time = 10,
   two_tailed = TRUE, # Select two-tailed tests for the post hoc tests (Default)
   posthoc_pairwise_tests = TRUE, # Ask for post hoc tests to be carried out
   rate_type = "net_diversification",
   seed = 1234, # Set seed for reproducibility
   return_perm_data = TRUE,
   verbose = FALSE)) # Keep permutation data to be able to plot histograms of tests

# Plot histogram of all post hoc two-tailed test stats
plot_histogram_STRAPP_test_for_focal_time(
  deepSTRAPP_outputs = deepSTRAPP_output_two_tailed,
  plot_posthoc_tests = TRUE)
```

```{r STRAPP_tests_cat_3lvl_one_tailed}
#### Step 4: Run all post hoc Dunn's one-tailed tests ####

# Here, we make assumptions on the direction of the effects tested.
# We conduct all possible one-tailed tests across pairs of states.
# Thus, for three pairs, we conduct six tests, one for each possible alternative hypothesis
# (i.e., "A" > "B" and "B" > "A").

# We run a one-tailed STRAPP test for t = 10 Mya.
focal_time <- 10

deepSTRAPP_output_one_tailed <- run_deepSTRAPP_for_focal_time(
   densityMaps = Ponerinae_trait_object$densityMaps,
   ace = Ponerinae_trait_object$ace,
   tip_data = Ponerinae_cat_3lvl_tip_data,
   trait_data_type = "categorical",
   BAMM_object = Ponerinae_BAMM_object_old_calib,
   focal_time = focal_time,
   two_tailed = FALSE, # Select one-tailed tests for the post hoc tests
   posthoc_pairwise_tests = TRUE, # Ask for post hoc tests to be carried out
   rate_type = "net_diversification",
   seed = 1234, # Set seed for reproducibility
   return_perm_data = TRUE) # Keep permutation data to be able to plot histograms of tests

## Explore output
str(deepSTRAPP_output_one_tailed, max.level = 2)

# Access deepSTRAPP results for post hoc tests
deepSTRAPP_output_one_tailed$STRAPP_results$posthoc_pairwise_tests$summary_df

# We obtain a p-value for each pair of states in each direction.
# Here there are 3 possible pairs = 6 one-tailed tests.
# Only the alternative hypothesis "arboreal < terricolous" yields to a significant result.
# For this test, we obtain a p-value = 0.012 leading us to discard the null hypothesis that 
# "arboreal" ants have equivalent or higher rates than "terricolous" ants.
# The estimate is the quantile 5% of the distribution of differences 
# in Dunn's Z-stats between observed and permuted data.
# The test is significant as the estimate is higher than zero (Q5% = 1.718).
# We can conclude that our data support the idea that "arboreal" ants
# have lower diversification rates than "terricolous" ants.

# Plot histogram of all post hoc one-tailed test stats
plot_histogram_STRAPP_test_for_focal_time(
  deepSTRAPP_outputs = deepSTRAPP_output_one_tailed,
  plot_posthoc_tests = TRUE)

# Only the test for the alternative hypothesis "arboreal < terricolous" is significant as the red line
# (quantile 5% = 1.718) is above the null expectation that Δ Dunn's Z-stats = 0.
# The conclusion is that it is the higher rates in "terricolous" ants vs. lower rates in "arboreal" ants
# that drives the differences in rates recorded overall with the Kruskal-Wallis test.
```
```{r STRAPP_tests_cat_3lvl_one_tailed_eval, fig.width = 8.5, fig.height = 6, out.width = "100%", eval = TRUE, echo = FALSE}
# We run a one-tailed STRAPP test for t = 10 Mya.
deepSTRAPP_output_one_tailed <- suppressWarnings(run_deepSTRAPP_for_focal_time(
   densityMaps = Ponerinae_trait_object$densityMaps,
   ace = Ponerinae_trait_object$ace,
   tip_data = Ponerinae_cat_3lvl_tip_data,
   trait_data_type = "categorical",
   BAMM_object = Ponerinae_BAMM_object_old_calib,
   focal_time = 10,
   two_tailed = FALSE, # Select one-tailed tests for the post hoc tests
   posthoc_pairwise_tests = TRUE, # Ask for post hoc tests to be carried out
   rate_type = "net_diversification",
   seed = 1234, # Set seed for reproducibility
   return_perm_data = TRUE, # Keep permutation data to be able to plot histograms of tests
   verbose = FALSE))

# Plot histogram of all post hoc one-tailed test stats
plot_histogram_STRAPP_test_for_focal_time(
  deepSTRAPP_outputs = deepSTRAPP_output_one_tailed,
  plot_posthoc_tests = TRUE)
```
